---
title: "613 Anaysis: Examining Life Expectancy"
author: "Luis Aguilar Levy"
date: "'r Sys.Date()'"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

Here, I re-re-examine the analysis of Oeppen and Vaupel (2002).

# Life Expectancy Analysis

We will try two full solutions


```{r}
library(tidyverse)
library(gapminder)
data(gapminder_unfiltered)
head(gapminder_unfiltered)
gapminder_unfiltered%>% 
  group_by(year)%>%
  summarize(maxlife = max(lifeExp, na.rm = TRUE)) -> cdf
head(cdf)
  
left_join(cdf,gapminder_unfiltered, by = c("maxlife"= "lifeExp"))
  

gapminder_unfiltered%>%
  group_by(year)%>%
  filter(lifeExp == max(lifeExp))%>%
  select(country, lifeExp, year)%>%
  arrange(year) -> sumdat

head(sumdat)
  


```

We will now recreate the plot from Oeppen and Vaupel(2002).



# 1st Class Notes
Other words for command line: shell, terminal, command line interface (cli), and console.

These terms are technically slightly different.

Bash: f x --y=1 #bash are functions
ls: list files command
pwd: shows which file you're at
cd: cd and .. up one level
linux subsytems
ls --help : help 'a' flag means all
cd
cd ~/../../.. : it took me back outside the computer or 6 levels
espcape spaces with a backslash '\'
use 'tab' to outocomplete file names
cp: copy file
touch: creates blank file of what you list

LAAL@DESKTOP-R7G4JGT MINGW64 ~/OneDrive/Documents/stat613/life_exp (master)
$ touch temp.txt

LAAL@DESKTOP-R7G4JGT MINGW64 ~/OneDrive/Documents/stat613/life_exp (master)
$ cp temp.txt temp_new.txt

rm: removes gile
mv: rename files 
mkdir: creates new directory
rmdir: remove directory

All git commands begin with git followed immediately by an argument for the type of command you want to execute.

LAAL@DESKTOP-R7G4JGT MINGW64 ~/OneDrive/Documents/stat613/life_exp (master)
$ git config --global user.name "laguilarl"

LAAL@DESKTOP-R7G4JGT MINGW64 ~/OneDrive/Documents/stat613/life_exp (master)
$ git config --global user.email "la2771a@student.american,edu"

1 Repo for each homework

git init
git add file.rmd
git status
git diff
Changes to be committed = staged files

git reset: everything from stage gets removed 
git commit -m "creates first rmd blank file"
wihtout "-m" you'll go into VIM hell.
Vim hell:
  i to insert in VIM
  escape + colon + wq to escape VIM

q to exit






## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
